### 2019-09-21 submission/manuscript.tex
work on results, add references

\documentclass[10pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.0in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Department Invited Speakers Do Not Reflect Trainee Diversity},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{\textbf{Department Invited Speakers Do Not Reflect Trainee Diversity}}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{}
    \preauthor{}\postauthor{}
    \date{}
    \predate{}\postdate{}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage[table]{xcolor}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{caption}

\usepackage{helvet} % Helvetica font
\renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font
\usepackage[T1]{fontenc}

\usepackage[none]{hyphenat}

\usepackage{setspace}
\doublespacing
\setlength{\parskip}{1em}

\usepackage{lineno}

\usepackage{pdfpages}
\floatplacement{figure}{H} % Keep the figure up top of the page

\begin{document}
\maketitle

\vspace{30mm}

Running title: Invited Speaker Diversity Does Not Reflect Trainee
Diversity

\vspace{35mm}

Ada K. Hagan, Ph.D.\({^1\dagger}\), Rebecca M. Pollet, Ph.D.\({^1}\),
and Josie Libertucci, Ph.D.\({^2\dagger}\)

\vspace{35mm}

\(\dagger\) To whom correspondence should be addressed:
\href{mailto:akhagan@umich.edu}{\nolinkurl{akhagan@umich.edu}} or
\href{mailto:libertj@mcmaster.ca}{\nolinkurl{libertj@mcmaster.ca}}

1. Department of Microbiology \& Immunology, University of Michigan, Ann
Arbor, Michigan

2. Department of Medicine, McMaster University, Hamilton, Ontario,
Canada

Figures: 1

Tables: 1

\newpage

\linenumbers

\subsection{Abstract}\label{abstract}

\subsection{Keywords}\label{keywords}

inclusion, diversity, invited speakers, academia, graduate programs

\newpage

\subsection{Background}\label{background}

Long-standing systemic bias, sexism, and racism have contributed to the
underrepresentation of many racial and ethnic groups, as well as women,
in science, technology, engineering, and math (STEM) fields (NIH 2015
report, Measuring DEI 2016, (1)). Organization climate and culture that
supports inclusion of all individuals is essential to support the
retention of historically underrepresented minorities (HURM) in the
U.S., as well as non-Caucasian individuals, and women in STEM fields
(Schneider, 2013). Inclusive organization culture will not suffice in
retaining HURM, non-Caucasian individuals, and women without
organization policy changes that support their values and goals (2).

A long-standing issue within STEM fields, specifically within academia,
has been a lack of diverse representation of scientists for trainees
(graduate students and postdoctoral fellows). In order to maintain
retention of HURM, non-Caucasian individuals, and women in STEM fields,
it is important for trainees to have visual representations of
themselves as scientists. The importance of representation in retaining
a diverse group of individuals in STEM fields is supported by social
role theory (3). Individuals tend to make inferences about
characteristics that are needed to be successful in a given role by
examining individuals that most occupy that role (3, 4). Therefore,
trainees who do not see representation of themselves in senior
scientific positions, or in this case as senior faculty members, may
decide that they do not possess the characteristics that are required to
succeed. In an attempt to make science a more inclusive environment,
many individuals have attempted to address this issue by promoting the
inclusion of more women speakers at conferences (5--7). However, no
study to date has addressed invited speaker seminar series diversity
within their own institution.

Within the Department of Microbiology and Immunology at the University
of Michigan (Ann Arbor, MI, USA), each year from September to June,
faculty members have the opportunity to invite scientists from other
institutions to give a one-hour seminar in which all department members
attend. The invited seminar speaker also has the opportunity to meet
with current department faculty members and trainees. Scientists who are
invited to give seminars are widely regarded as successful and the top
in their field. Thus, if trainees are constantly being exposed to ``the
top scientist in their field'', according to social role theory, it is
imperative that the selected seminar speakers represent a diverse group
of individuals.

In this study, we examine and compare the proportion of HURM,
non-Caucasian/non-HURM, and women invited speakers to white males in the
Department of Microbiology and Immunology. Additionally, we compare
invited-speaker demographics to the current trainee demographics as a
means to gauge if trainee demographics are being represented accordingly
throughout the seminar series. Following our investigation, we proposed
a policy change to the Department of Microbiology and Immunology in how
invited speakers are selected as a means to promote inclusion in our
department and reduce stereotype threat, microaggressions, and
unconscious bias. In order to facilitate inviting a more diverse group
of scientists, we developed a set of resources that allow scientists,
within the field of microbiology, to self-identify as an HURM,
non-Caucasian/non-HURM, or a woman. These resources will promote
inclusion and diversity by providing greater representation of all
scientists and will provide hosts an opportunity to invite a more
diverse group of scientists.

\subsection{Methods}\label{methods}

Each academic year, each faculty member in the Department of
Microbiology and Immunology at the University of Michigan has the
opportunity to invite one speaker per year for a weekly seminar series.
Some of these seminar slots are dedicated to named lectureships, which
are decided by committee, and three trainee-invited speakers. We
analyzed the demographics of invited speakers and faculty hosts for five
academic years (Fall 2014 - Spring 2019), and compared them to the
current trainees when the data were analyzed (Spring 2019). Each speaker
was only counted once and those listed as departmental faculty members
or as a ``host'' at any point could not also be considered ``invited
speakers''. The list of faculty hosts was used as a proxy for faculty
demographics since as hosts, these faculty members are visible
representatives of the department. The trainee lists were obtained from
department listservs that included masters students, doctoral students,
and post-doctoral fellows.

This is a retrospective study, so individuals were not asked for their
identities. Instead we hand-coded proxy demographics using personal
knowledge, photos, and CVs. The presenting gender of each individual was
assigned using a binary system (man/woman). Due to the low number of
individuals in the study, race/ethnicity demographics were split in
three groups: Caucasian, Historically Under-represented Minority (HURM),
and Non-Caucasian/Non-HURM (NCNH), each with a binary (yes/no)
possibilty. Caucasian was assigned using the current U.S. Census
definition where those of Middle Eastern, European, and Russian descent
are included. HURM individuals were restricted to those with
African-American, Indigenous, Alaskan/Hawaian Native, Latinx and/or
Hispanic heritage. All others were placed into the NCNH group. We
recognize that our proxy demographics are a limitation of the analysis
and want to acknowledge that the presenting gender may not reflect the
gender(s) that an individual identifies as. We also want to acknowledge
that there are many other identities that are not captured in this
limited analysis.

Data were analyzed and figures generated in R Statisical Software, using
relevant packages (8--20).

\subsection{Results}\label{results}

To understand the representation of women, we compared the proportion of
women in each academic role. At the trainee level, more than half of
students and postdoctoral fellows were women. That dropped to 46.77\% of
faculty hosts and 38.73\% of the invited speakers (Fig. 1A). Of 27
lectureships over the five year period, 37.04\% were awarded to women.
The proportion of women as faculty hosts and speakers is equivalent to
global estimates that 40\% of microbiologists are women, with a slightly
lower representation of women in lectureships (21).

Our analysis identified an over-representation of Caucasian individuals
as hosting faculty and invited speakers (80\% each), relative to the
proportion of Caucasian trainees, which was 55\% (Fig. 1B). We also
observed declines in the representation of HURM and NCNH faculty and
speakers relative to the trainees (Fig 1B). HURM trainees made up 11\%
of the department, on track with the 11\% of microbiology and immunology
doctorates awarded in 2017 (22). However, only 8.5\% of invited
speakers, and none of the hosting faculty, were HURM scientists. NCNH
trainees were 34\% of department students and postdocs (22\% of
microbiology and immunology doctorates in 2017), but only 19\% of
hosting faculty and 10.5\% of invited speakers (22).

The more prestigous invited speaker lectureships were also dominated by
caucasian scientists, who comprised 81.48\% of those awarded (Fig. 1C).
HURM and NCNH scientists were awarded 3 and 2 lectureships,
respectively. Because the intersection of identities can compound biases
and outcomes, we further examined the lectureships by gender and
race/ethnicity status. Caucasian men and women accounted for 44.44\% and
37.04\% of the lectureships, respectively. Just 18.52\% of lectureships
were held by non-Caucasian men while none were held by non-Caucasian
women (Fig. 1D).

\subsection{Discussion}\label{discussion}

Several papers have investigated the representation of women at
scientific conferences, however, we have only identified one that
focused on invited speakers at universities (23). In their study,
Nittrouer et al, examined 3,652 talks at 50 U.S. institutions in 2013 -
2014 and found that women faculty are less likely to be invited
speakers, despite similar acceptance rates. These results suggest that
women faculty are less often invited as speakers, a decision that may be
negatively impacted by assumptions about competency and dedication. The
dedication of women who have children to their work is perceived to be
less than that of their colleagues, i.e., men who also have children.
The perceived prioritization and commitments of women to family over
work may cause faculty to doubt their acceptance of a speaking
invitation (despite the prestigious nature of these invitations and
evidence to the contrary)({\textbf{???}}, 23). As a result, the faculty
member invites a different colleague who they feel is more likely to
agree (and is a man). Departments have different processes and criteria
for selecting invited speakers, but it is a matter of pride to bring the
best scientists possible. It may be that the definition of ``best''
poses a problem to women, who need three-times as many publications as
their men colleges to be considered equally competent. Some departments
only invite tenured faculty, which severely limits the number of
potential women speakers. Another scenario is that pre-tenure faculty
members invite prestigious, tenured faculty in their field to network
and secure letters for their own tenure package. The increased burden of
women to prove competency decreases their likelihood to be considered
for either tenure or as possible source of tenure letters.

The desire to invite the ``best'' scientist to these seminar series also
poses a problem for HURM and NCNH scientists. A large portion of the
NCNH cohort in our sample are Asian/Asian American individuals. Although
Asian scientists are well-represented in the US scientific workforce,
they face significant bias and barriers to inclusion in society and
academia (24, 25). Both men and women who identify as HURM and Asian
reported having to work harder than white men to be perceived as
legitimate scholars (26). This expectation of higher productivity may
result in fewer HURM and NCNH individuals being invited as seminar
speakers. This is particularly striking in the low number of NCNH
individuals in the more prestigious lectureships of our sample despite
the higher employment rate of Asian scientists.

While HURM and NCNH share some experiences, differences including
varying rates of hiring and tenure promotion mean unique considerations
are important for inclusion of each group
(\url{doi:10.1001/jama.284.9.1085}). For instance, the proportion of
HURM faculty at the Assistant and Associate Professor level is currently
higher than the Full Professor level so it will be difficult to increase
speaker diversity if early-career researchers are not being considered
(27). Increased performance expectations and patterns of exclusions are
consistent themes in studies characterizing the HURM faculty experience
(28, 29). Therefore, inclusion of HURM faculty in seminar series is
likely essential to increasing the number of HURM Associate and Full
Professors. More significantly, a major barrier to inclusion of HURM
faculty at similar proportions to HURM trainees is the low transition
rate of scientists from HURM backgrounds to faculty positions and the
associated low proportion of HURM faculty (30). Even when HURM speaker
rates match the proportion of HURM faculty employment, HURM trainees
will be represented at a significantly higher proportion. Inclusion of
HURM faculty in these seminarseries is just one aspect of larger
institutional change that is needed (31).

We have not been able to identify any other publications examining
scientific speaker diversity beyond gender. This seems to be the first.
This is concerning since conclusions drawn from gender-based studies are
often framed, and considered, to be applicable to other marginalized
groups (e.g., HURM). This is a flawed assumption. While there is no
doubt some overlap, each group remains marginalized due to a unique
complex set of factors that cannot always be solved by gender-based
solutions. US-serving institutions, such as the University of Michigan
have a particular responsibility to the historically suppressed
populations included in our definition of HURMs. We therefore implore US
institutions to apply this framing to their discussions and research.

According to social role theory, this department is not adequietly
serving their non-Caucasian trainees. Implict biases that affect
perceptions of marginalized groups are the primary issue, but we must
acknowledge that it is not always possible to identify members of
historically under-served communities. For instance, after data
analysis, we learned that at least one speaker in our data set should
have been categorized as a HURM instead of Caucasian, but it wasn't
readily apparent from their internet presence or CV. This limitation
makes two important points: that percieved identity often plays a larger
role than self-identification, and that we need better tools to identify
members of marginalized groups.

\subsection{Instituting policy change within the Department of
Microbiology and
Immunology}\label{instituting-policy-change-within-the-department-of-microbiology-and-immunology}

In an attempt to promote inclusion within the Dept. M\&I these data were
presented to faculty members and the dept chair (Harry Mobley). Since,
trainee demographics were not represented within seminar speaker
demographics over the past 5 years, we proposed a policy change as to
how seminar speakers were being invited. This policy change included
switching from PI invited to lab invited in an attempt to allow trainees
to choose a speaker that best represented themselves. In addition to
this we also created resources which allowed scientists to self-identify
and provide host faculty with more diverse choices.

\subsection{Building Diversify}\label{building-diversify}

To help address this issue, we make some suggestions (Table 1) and have
developed a resource to identify scientists who are members of
marginalized and/or historically under-served groups. Motivated by a
lack of such resources and inspired by similar resources--DiversifyEEB
and DiversifyChemistry--we created DiversifyMicrobiology and
DiversifyImmunology. These resources are a tool for symposium
organziers, award committees, search committees and other scientists to
identify individuals to diversify their pools. Additionally, we have
built these as a tool for use by other fields and organizations to
create their own lists. Importantly, since these lists are compiled by
self-nomination, we can ensure that only scientists comfortable
revealing their marginalized identities are included.

The self-nomination form is a Google Form with entries logged in a
private Google Sheet. This form is embedded within the website and can
be linked to directly. The use of a Google Forms allows us to maintain
this database at no cost and gives us the flexibility to add questions
or change response options without disrupting previous responses.
Entries are logged in a private spreadsheet so that entries can be
screened before being added to the public database. This screening
currently includes two steps: 1. Confirming that each person is listed
in the database only once. 2. Verifying that the submitted website is a
professional website for the person being listed. If a person has not
been listed in the database previously and passes step 2, a new entry is
added to the public database spreadsheet. If a person is already listed
in the database, their information is updated to the most recent
submission.

This public spreadsheet is embedded in the website and can be open
separately as a locked (uneditable) Google Sheet. Importantly, this
allows the list to be easily searched. We have chosen to list
individuals' academic information first in the spreadsheet to encourage
a focus on academic achievement rather than tokenization of marginalized
identities. Currently the database lists individuals in order of
self-nomination but future versions will be re-sorted based on name
and/or academic field to varying the individuals who may receive more
attention for simply being at the top of the list.

The website provides an interface to the Google forms and spreadsheets
with template pages for viewing the list, adding a name to the list, and
finding additional resources. Importantly, our website creation tool is
hosted for free by GitHub, which provides a free website for each GitHub
organization. Basic tools and skills required to set up a Diversify site
include knowledge of, or experience with, the version control tool git,
the webtool GitHub, and a text editor. A tutorial in the
DiversifyMicrobiology repository on GitHub provides links to these
resources and instructions for adapting the tool to your own field.

\subsection{Conclusion}\label{conclusion}

To increase the retention of white women and HURMS in STEM, they must
also be represented as experts. However, the invited speaker diversity
at one department does not represent the diversity of trainees. There is
a lack of research on invited speakers examining factors other than
gender. To facilite the identification and recruitment of individuals in
these historically under-served groups, we have built a tool to create
self-nominated, field-specific lists.

\subsection{Acknowledgements}\label{acknowledgements}

We thank Drs. Beth Moore and Harry Mobley and the Department of
Microbiology \& Immunology, University of Michigan for their input and
financial support that enabled publication of our manuscript. We would
also like to acknowledge Nick Lesniak and Dr.~Ariangela Kozick for their
comments and suggestions.

\subsection{Author Contributions}\label{author-contributions}

A.K.H. collected the data, assigned demographics, analyzed the data, and
created the website. R.M.P. created the Google lists, forms, and website
content and the description of their maintenance. J.L. wrote the
introduction and provided conceptual advice. All authors contributed to
preparing the final manuscript.

\subsection{Code and data
availability}\label{code-and-data-availability}

The anonymized data, code for all analysis steps, and an Rmarkdown
version of this manuscript is available at
\url{https://github.com/akhagan/Hagan_SpeakerDiversity_XXXX_2019/}.
Template and complete instructions for generating a field-specific
Diversity website are available at
\url{https://github.com/diversifymicrobiology/DiversifyMicrobiology.github.io/}.

\begin{figure}
\centering
\includegraphics{Figure_1.png}
\caption{\textbf{The demographics of invited speakers, hosting faculty,
and trainees.} A) The proportion of women in each academic role. B) The
proportion of each academic role represented by individuals that are
Caucasian (left), Historically Underrepresented Minorities (HURM,
center) or International (right). C-D)The percent of lectureships
awarded to individuals that are C) Caucasian, HURM, International or D)
Caucasian or Non-Caucasian by gender.}
\end{figure}

\newpage

\begin{center}
\captionof{table}{List of suggestions and resources to increase invited speaker diversity.}
\small
\begin{tabular}{|l|l|l|}
\hline

\rowcolor{lightgray}
\textbf{Suggestion} & \textbf{Description} & \textbf{Resource} \\ \hline

Lab-invited speakers & \makecell[l]{Faculty members can request \\suggestions from trainees} & \\ \hline

Use a list & \makecell[l]{Many lists of scientists from \\under-represented and underserved \\groups are available} &  \makecell[l]{https://DiversifyMicrobiology.\\github.io/resources}\\ \hline

Create a list & \makecell[l]{Use the GitHub template \\ create a self-nomination list and \\resource for your field} & \makecell[l]{https://github.com/diversifymicrobiology/\\DiversifyMicrobiology.github.io} \\ \hline

Highlight the journey & \makecell[l]{Invite all speakers to spend \\a few moments describing their \\personal science journey} & \\ \hline

\end{tabular}
\end{center}

\newpage

\subsection{References}\label{references}

\hypertarget{refs}{}
\hypertarget{ref-colloquy_2012}{}
1. 2012. Colloquy on minority males in science, technology, engineering,
and mathematics. National Academies Press.

\hypertarget{ref-Coe2019}{}
2. \textbf{Coe IR}, \textbf{Wiley R}, \textbf{Bekker L-G}. 2019.
Organisational best practices towards gender equality in science and
medicine. The Lancet \textbf{393}:587--593.
doi:\href{https://doi.org/10.1016/s0140-6736(18)33188-x}{10.1016/s0140-6736(18)33188-x}.

\hypertarget{ref-Eagly1984}{}
3. \textbf{Eagly AH}, \textbf{Steffen VJ}. 1984. Gender stereotypes stem
from the distribution of women and men into social roles. Journal of
Personality and Social Psychology \textbf{46}:735--754.
doi:\href{https://doi.org/10.1037/0022-3514.46.4.735}{10.1037/0022-3514.46.4.735}.

\hypertarget{ref-Carter2018}{}
4. \textbf{Carter AJ}, \textbf{Croft A}, \textbf{Lukas D},
\textbf{Sandstrom GM}. 2018. Women's visibility in academic seminars:
Women ask fewer questions than men. PLOS ONE \textbf{13}:e0202743.
doi:\href{https://doi.org/10.1371/journal.pone.0202743}{10.1371/journal.pone.0202743}.

\hypertarget{ref-kalejta_gender_2017}{}
5. \textbf{Kalejta RF}, \textbf{Palmenberg AC}. 2017. Gender Parity
Trends for Invited Speakers at Four Prominent Virology Conference
Series. Journal of Virology \textbf{91}.
doi:\href{https://doi.org/10.1128/JVI.00739-17}{10.1128/JVI.00739-17}.

\hypertarget{ref-casadevall_presence_2014}{}
6. \textbf{Casadevall A}, \textbf{Handelsman J}. 2014. The Presence of
Female Conveners Correlates with a Higher Proportion of Female Speakers
at Scientific Symposia. mBio \textbf{5}.
doi:\href{https://doi.org/10.1128/mBio.00846-13}{10.1128/mBio.00846-13}.

\hypertarget{ref-klein_speaking_2017}{}
7. \textbf{Klein RS}, \textbf{Voskuhl R}, \textbf{Segal BM},
\textbf{Dittel BN}, \textbf{Lane TE}, \textbf{Bethea JR}, \textbf{Carson
MJ}, \textbf{Colton C}, \textbf{Rosi S}, \textbf{Anderson A},
\textbf{Piccio L}, \textbf{Goverman JM}, \textbf{Benveniste EN},
\textbf{Brown MA}, \textbf{Tiwari-Woodruff SK}, \textbf{Harris TH},
\textbf{Cross AH}. 2017. Speaking out about gender imbalance in invited
speakers improves diversity. Nature Immunology \textbf{18}:475--478.
doi:\href{https://doi.org/10.1038/ni.3707}{10.1038/ni.3707}.

\hypertarget{ref-R_software_2017}{}
8. \textbf{R Core Team}. 2017. R: A language and environment for
statistical computing. R Foundation for Statistical Computing, Vienna,
Austria.

\hypertarget{ref-wickham_tidyverse_2017}{}
9. \textbf{Wickham H}. 2017. Tidyverse: Easily Install and Load the
'Tidyverse'.

\hypertarget{ref-cowplot}{}
10. \textbf{Wilke CO}. 2019. Cowplot: Streamlined plot theme and plot
annotations for 'ggplot2'.

\hypertarget{ref-markdown}{}
11. \textbf{Allaire J}, \textbf{Horner J}, \textbf{Xie Y}, \textbf{Marti
V}, \textbf{Porte N}. 2018. Markdown: 'Markdown' rendering for r.

\hypertarget{ref-rmd_book}{}
12. \textbf{Xie Y}, \textbf{Allaire J}, \textbf{Grolemund G}. 2018. R
markdown: The definitive guide. Chapman; Hall/CRC, Boca Raton, Florida.

\hypertarget{ref-rmd_rstudio}{}
13. \textbf{Allaire J}, \textbf{Xie Y}, \textbf{McPherson J},
\textbf{Luraschi J}, \textbf{Ushey K}, \textbf{Atkins A},
\textbf{Wickham H}, \textbf{Cheng J}, \textbf{Chang W}, \textbf{Iannone
R}. 2018. Rmarkdown: Dynamic documents for r.

\hypertarget{ref-knitr_2014}{}
14. \textbf{Xie Y}. 2014. Knitr: A comprehensive tool for reproducible
research in R. \emph{In} Stodden, V, Leisch, F, Peng, RD (eds.),
Implementing reproducible computational research. Chapman; Hall/CRC.

\hypertarget{ref-knitr_2018}{}
15. \textbf{Xie Y}. 2018. Knitr: A general-purpose package for dynamic
report generation in r.

\hypertarget{ref-lubridate}{}
16. \textbf{Grolemund G}, \textbf{Wickham H}. 2011. Dates and times made
easy with lubridate. Journal of Statistical Software \textbf{40}:1--25.

\hypertarget{ref-readxl}{}
17. \textbf{Wickham H}, \textbf{Bryan J}. 2018. Readxl: Read excel
files.

\hypertarget{ref-pdftools}{}
18. \textbf{Ooms J}. 2019. Pdftools: Text extraction, rendering and
converting of pdf documents.

\hypertarget{ref-wickham_scales_2018}{}
19. \textbf{Wickham H}. 2018. Scales: Scale Functions for Visualization.

\hypertarget{ref-neuwirth_rcolorbrewer_2014}{}
20. \textbf{Neuwirth E}. 2014. RColorBrewer: ColorBrewer Palettes.

\hypertarget{ref-allagnat_gender_2017}{}
21. \textbf{Allagnat L}, \textbf{Berghmans S}, \textbf{Falk-Krzesinski
HJ}, \textbf{Hanafi S}, \textbf{Herbert R}, \textbf{Huggett S},
\textbf{Tobin S}. 2017. Gender in the global research landscape.

\hypertarget{ref-nsf_survey_2017}{}
22. \textbf{Science and Engineering Statistics NC for}. 2017. Survey of
Doctorate Recipients, Survey Year 2017. National Science Foundation,
Alexandria, VA.

\hypertarget{ref-nittrouer_gender_2018}{}
23. \textbf{Nittrouer CL}, \textbf{Hebl MR}, \textbf{Ashburn-Nardo L},
\textbf{Trump-Steele RCE}, \textbf{Lane DM}, \textbf{Valian V}. 2018.
Gender disparities in colloquium speakers at top universities.
Proceedings of the National Academy of Sciences \textbf{115}:104--108.
doi:\href{https://doi.org/10.1073/pnas.1708414115}{10.1073/pnas.1708414115}.

\hypertarget{ref-Hwang2008}{}
24. \textbf{Hwang W-C}, \textbf{Goto S}. 2008. The impact of perceived
racial discrimination on the mental health of asian american and latino
college students. Cultural Diversity and Ethnic Minority Psychology
\textbf{14}:326--335.
doi:\href{https://doi.org/10.1037/1099-9809.14.4.326}{10.1037/1099-9809.14.4.326}.

\hypertarget{ref-Tran2019}{}
25. \textbf{Tran VC}, \textbf{Lee J}, \textbf{Huang TJ}. 2019.
Revisiting the asian second-generation advantage. Ethnic and Racial
Studies \textbf{42}:2248--2269.
doi:\href{https://doi.org/10.1080/01419870.2019.1579920}{10.1080/01419870.2019.1579920}.

\hypertarget{ref-noauthor_seeking_2013}{}
26. 2013. Seeking Solutions: Maximizing American Talent by Advancing
Women of Color in Academia: Summary of a Conference. National Academies
Press, Washington, D.C.

\hypertarget{ref-whittaker_retention_2015}{}
27. \textbf{Whittaker JA}, \textbf{Montgomery BL}, \textbf{Martinez
Acosta VG}. 2015. Retention of Underrepresented Minority Faculty:
Strategic Initiatives for Institutional Value Proposition Based on
Perspectives from a Range of Academic Institutions. Journal of
undergraduate neuroscience education: JUNE: a publication of FUN,
Faculty for Undergraduate Neuroscience \textbf{13}:A136--145.

\hypertarget{ref-pololi_race_2010}{}
28. \textbf{Pololi L}, \textbf{Cooper LA}, \textbf{Carr P}. 2010. Race,
Disadvantage and Faculty Experiences in Academic Medicine. Journal of
General Internal Medicine \textbf{25}:1363--1369.
doi:\href{https://doi.org/10.1007/s11606-010-1478-7}{10.1007/s11606-010-1478-7}.

\hypertarget{ref-hassouneh_experiences_2014}{}
29. \textbf{Hassouneh D}, \textbf{Lutz KF}, \textbf{Beckett AK},
\textbf{Junkins EP}, \textbf{Horton LL}. 2014. The experiences of
underrepresented minority faculty in schools of medicine. Medical
Education Online \textbf{19}:24768.
doi:\href{https://doi.org/10.3402/meo.v19.24768}{10.3402/meo.v19.24768}.

\hypertarget{ref-Gibbs2016}{}
30. \textbf{Gibbs KD}, \textbf{Basson J}, \textbf{Xierali IM},
\textbf{Broniatowski DA}. 2016. Decoupling of the minority PhD talent
pool and assistant professor hiring in medical school basic science
departments in the US. eLife \textbf{5}.
doi:\href{https://doi.org/10.7554/elife.21393}{10.7554/elife.21393}.

\hypertarget{ref-johnson_msphere_2019}{}
31. \textbf{Johnson MDL}. 2019. mSphere of Influence: Hiring of
Underrepresented Minority Assistant Professors in Medical School Basic
Science Departments Has a Long Way To Go. mSphere \textbf{4}.
doi:\href{https://doi.org/10.1128/mSphere.00599-19}{10.1128/mSphere.00599-19}.

\end{document}